Online Classification of arXiv Papers Based on Your Preferences

### 2014-05-16 Paper.tex
edited paper.

\documentclass[12pt]{article}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{graphicx}
\pagestyle{myheadings}
\markright{Joseph McDonald}

@@

In Section 3, we describe the main components of the program in the python modules we created, and give a description on how to use the tool. We present our results in Section 4, followed by other observations and concluding remarks.

@@

Let $I_1$ be the set of $t$ where a mistake was made and
$\widehat{y}=1$ (false positives).  Let $I_2$ be the set where a mistake was made with
$\widehat{y} = -1$ and $k = 1$ (false negatives where the algorithm updates), and $I_3$ the set where a mistake was
made with $\widehat{y} = -1$ and $k = 0$ (false negatives where the algorithm does {\it not} make an update). Finally, let $M_i = |I_i|$ for $i =
1,2,3$. We bound $M_1 + M_2$ in the same way we bound $M$ in the original Perceptron algorithm. Assume there exists a $\rho$ and $R$ satisfying the assumptions of Theorem 7.8 from our text, and a separating hyperplane determined by the vector $v$. Let $M' = M_1 + M_2$, then
\begin{align*}
M' \rho =&\ \frac{v}{\|v\|}\sum_{I_1 \cup I_2} y_tx_t \
 \leq   \Bigl\|\sum_{t \in I_1 \cup I_2} y_t x_t\Bigr\| \\
=& \ \Bigl\|\sum_{t \in I_1 \cup I_2} (w_{t+1} - w_t) \Bigr\| 
=  \|w_{T+1} \| \\
=&\ \sqrt{\sum_{t \in I_1 \cup I_2} \|  w_{t+1}\|^2 - \|w_t\|^2 }\\
= &\ \sqrt{\sum_{t \in I_1 \cup I_2} \|  w_{t} + y_t x_t\|^2 - \|w_t\|^2 }\\
= &\ \sqrt{\sum_{t \in I_1 \cup I_2} 2y_t w_t\cdot x_t - \|x_t\|^2} \\
\leq &\ \sqrt{\sum_{t \in I_1 \cup I_2} \|x_t\|^2}
\leq \sqrt{M'r^2}.
\end{align*}
This implies that $M' \leq \frac{r^2}{\rho^2}$. Since $M = M' + M_3$, a bound on $M_3$ is needed. We revert to a weaker bound and consider the expectation of
$M_3$. Since 
\[M_3 = \sum_{t=1}^T 1_{\left\{\widehat{y}_t = -1 \cap y_t = 1\right\}}1_{k_t = 0}, \]
we get that
\begin{align*}
\E[M_3] = & \sum_{t=1}^T \E[1_{\left\{\widehat{y_t} = -1 \cap y_t = 1\right\}}]\E[1_{k_t = 0}] \\
 = &\ (1 - p)\sum_{t=1}^T \E[1_{\left\{\widehat{y_t} = -1 \cap y_t = 1\right\}}],
\end{align*}
where the last equality uses the independence of $k_t$ from the data. To control the sum of expectations in this expression we consider the expectation of $M_2$. Since
\[
M_2 =  \sum_{t=1}^T 1_{\left\{\hat{y_t} = -1 \cap y_t = 1\right\}}1_{k_t = 1},
\]
then
\begin{align*}
\E[M_2] =&\ \sum_{t=1}^T \E[1_{\left\{\hat{y_t} = -1 \cap y_t = 1\right\}}]\E[1_{k_t = 1}] \\
=&\ p\sum_{t=1}^T \E[1_{\left\{\hat{y_t} = -1 \cap y_t = 1\right\}}].
\end{align*}
Thus
\[
E[M_3] = \ \frac{1-p}{p}\ \E[M_2].
\]
We can apply the crude bound $\E[M_3] \leq \frac{1-p}{p}\ \E[M']$, since $M_2 \leq M'$. Thus the expected number of mistakes can be bounded by
\begin{align*}
\E[M]  =  &\ \E[M'] + \E[M_3] \\
\leq & \ \frac{r^2}{\rho^2} + \frac{1-p}{p}\frac{r^2}{\rho^2} \\
\leq  & \ \frac{1}{p}\frac{r^2}{\rho^2}
\end{align*}
Note that this is a weak bound.  Not only is it greater than the
original perceptron bound by a factor of $\frac{1}{p}$, it is also only a
bound in expectation.

@@

\noindent Adaptive $p$ Perceptron algorithm:
\begin{quotation}
%\noindent Modified Perceptron:\\
\noindent $w_1=0$\\
\noindent $p=1$\\
{\tt for $t=1$ to $T$}:\\
\indent receive($x_t$)\\
\indent $\widehat{y}_t=\mbox{sign}(w_t\cdot x_t)$\\
\indent {\tt if} $\widehat{y}_t =1$ : \\
\indent\indent receive($y_t$)\\
\indent\indent {\tt if $y_t\neq\widehat{y}_t$}:\\
\indent\indent\indent $w_{t+1} = w_t + y_t x_t$\\
\indent\indent {\tt else $w_{t+1} = w_t$}\\
\indent {\tt else if ($\widehat{y}_t =-1$ {\tt and} $k_t = 1$)}:\\
\indent\indent {\tt if $y_t\neq\widehat{y}_t$}:\\
\indent\indent\indent $w_{t+1} = w_t + y_t x_t$\\
\indent\indent\indent $p=p + \gamma_1(1 - p)$\\
\indent\indent {\tt else }\\
\indent\indent\indent $w_{t+1} = w_t$\\
\indent\indent\indent $p = p - \gamma_2(p - p_{\min})$\\
\indent {\tt else if $\widehat{y}_t = -1$ and $k_t=0$}:\\
\indent\indent $w_{t+1} = w_t$\\
{\tt return $w_{T+1}$}
\end{quotation}
Note that $p_{\min}$, $\gamma_1$ and $\gamma_2$ are parameters that can be tuned. It is easy to extend the bound above from the modified Perceptron to the adaptive $p$ case.  Let $I_i$, $M_i$, and $M'$ be defined as above. The bound $M'\leq r^2/\rho^2$ follows in exactly the same way.  We then bound $M_3$ in expectation, 
\begin{align*}
\E[M_3] = & \sum_{t=1}^T \E[1_{\left\{\widehat{y_t} = -1 \cap y_t = 1\right\}}]\E[1_{k_t = 0}] \\
 \leq &\ (1 - p_{\min})\sum_{t=1}^T \E[1_{\left\{\widehat{y_t} = -1 \cap y_t = 1\right\}}],
\end{align*}
and bound $\E[M_2]$ from below
\begin{align*}
\E[M_2] =&\ \sum_{t=1}^T \E[1_{\left\{\hat{y_t} = -1 \cap y_t = 1\right\}}]\E[1_{k_t = 1}] \\
\geq &\ p\sum_{t=1}^T \E[1_{\left\{\hat{y_t} = -1 \cap y_t = 1\right\}}].
\end{align*}
Continuing as before, this gives the following bound on $\E[M]$,
\begin{align*}
\E[M] = &\ \E[M'] + \E[M_3] \leq \frac{r^2}{\rho^2} + \frac{1-p_{\min}}{p_{\min}} \frac{r^2}{\rho^2}\\
= &\ \frac{1}{p_{\min}}\frac{r^2}{\rho^2}
\end{align*}
Unfortunately, we do not have a stonger theoretical bound for this adaptive $p$ approach, but we test this scheme in section \ref{sec:results} and see that it performs fairly well while requiring less input from the user.

@@

The files {\tt Perceptron.py} and {\tt OnlineSVM.py} house the classes that compose the different classifying algorithms we use.
The classes contain the methods that run the algorithm and the attributes such as weights and parameters that the algorithm uses and updates.
Within {\tt Perceptron.py} is the modified Perceptron algorithm we presented earlier, a Margin Perceptron classifier, and a Kernel Perceptron classifier supplied with both a polynomial and Gaussian kernel.
Similarly, the Online SVM algorithm is implemented in the file {\tt OnlineSVM.py}.
Each algorithm takes a learning rate $\eta$ as a parameter. The modified Perceptron requires no extra parameters, while the Gaussian Kernel Perceptron classifier requires a $\sigma$ specifying the variance in the kernel.
The Polynomial Kernel Perceptron takes two parameters $c$ and $d$ which are the constants used in constructing the polynomial kernel.
The Margin Perceptron requires a parameter $\rho$ designating the separating margin of the feature vectors, and the Online SVM algorithm takes a parameter $C$ for the penalty applied to the update (i.e. $w_{t+1} = w_t -\eta(w_t- Cy_tx_t)$).
Each algorithm class has an associated wrapper that instantiates a class, and we use each of these algorithms when testing the tool.

@@

We tested the different online learning algorithms with sample data generated by graduate students at NYU.  Several students were asked to use {\tt getTestData.py} and {\tt LabelData.py} to gather the abstracts from Arxiv in categories that they specified (see SECTION REFERENCE HERE) and generate labels for each entry indicating whether they would be interested in reading a paper based on the abstract.

@@

It was interesting to see how varying $p$ the probability of presenting the paper had an impact on the results, some which might not have been entirely expected. For instance I would expect that asking for the most labels from the user (i.e. $p = 1$) would translate to considerably higher gains in both precision and recall all around. Yet for the best performing algorithms there wasn't a huge difference in recall between the adaptive version and $p = 1$, and for precision both adaptive $p$ and $p = 0.1$ outperformed $p = 1$. In fact this was generally true for all algorithms.

\begin{figure}\centering
\includegraphics[width=0.7\textwidth]{EveryoneRecallPaper.pdf}
\includegraphics[width=0.7\textwidth]{EveryonePrecisionPaper.pdf}
\label{fig:PrecisionandRecall}
\caption{Performance of tested online classification algorithms using different values of $p$.  The adaptive $p$ algorithm performs relatively well compared to the original algorithm ($p = 1$), and requires less than half as many negative updates.  Majority vote classifier combines all others in this figure.}
\end{figure}

\section{Conclusions}
To summarize our work, we're pleased to have a developed a tool that we believe can be of use for many people across several fields and to have employed a number of machine learning features in its construction, from the text categorization model to the classification algorithms. Beyond constructing these features we created some of our own features. For instance the adaptive probability $p$ turned out to perform fairly well in comparison to the tool that didn't filter results based on preferences. It was also interesting to see the bound that we got adapted from the original Perceptron model. One point to mention about future work in this area is that it seems likely that some of the other bounds from the Perceptron algorithm can probably be applied to get similar bounds for this problem as well.

Online learning methods are used to learn the user's preferences over time, but in order for them to be useful, they cannot present the user with every single prediction.  Instead, we weaken the algorithms to present the user with all articles it predicts will be interesting, and only some of those it feels will not be interesting.  We show that in this context, a bound on the expected number of mistakes can still be achieved under the some assumptions.  Finally, we test several algorithms on labeled data, and show that they perform comparably to the original learning algorithms, but require less input from the user.

In the future, it would be nice to develop tighter theoretical bounds on our modified learning algorithm, and particularly to develop an improved bound on the adaptive $p$ algorithm.  It is possible that such a bound would lead to a more informed decision on how to change $p$ in the algorithm to minimize mistakes. Perhaps giving a different functional form to the way $p$ is changed can produce an algorithm that minimizes updates, though this guess is comes from our experience with the AdaBoost algorithm and its modifications.

Finally, it would be worthwhile to explore other online learning algorithms and possibly other features that might improve predictive performance. Considering a wider variety of text classification models would probably produce better results, and the limited time to develop this tool compromised our ability to develop and play around with some of these models. Nevertheless that is a huge body of work in and of itself.
\end{document}

@@

>>>>

\documentclass[12pt]{article}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{graphicx}
\usepackage{natbib}
\pagestyle{myheadings}
\markright{Joseph McDonald}

@@

In Section 3, we describe the main components of the program in the python modules we created, and give a description on how to use the tool. We present our results in Section 4, followed by other observations and concluding remarks in seciont 5.

@@

Let $I_1$ be the set of $t$ where a mistake was made and
$\widehat{y}=1$ (false positives).  Let $I_2$ be the set where a mistake was made with
$\widehat{y} = -1$ and $k = 1$ (false negatives where the algorithm updates), and $I_3$ the set where a mistake was
made with $\widehat{y} = -1$ and $k = 0$ (false negatives where the algorithm does {\it not} make an update). Finally, let $M_i = |I_i|$ for $i =
1,2,3$. We bound $M_1 + M_2$ in the same way we bound $M$ in the original Perceptron algorithm. Assume there exists a $\rho$ and $R$ satisfying the assumptions of Theorem 7.8 from \cite{FML}, and a separating hyperplane determined by the vector $v$. Let $M' = M_1 + M_2$, then
\begin{align*}
M' \rho =&\ \frac{v}{\|v\|}\sum_{I_1 \cup I_2} y_tx_t \
 \leq   \Bigl\|\sum_{t \in I_1 \cup I_2} y_t x_t\Bigr\| \\
=& \ \Bigl\|\sum_{t \in I_1 \cup I_2} (w_{t+1} - w_t) \Bigr\| 
=  \|w_{T+1} \| \\
=&\ \sqrt{\sum_{t \in I_1 \cup I_2} \|  w_{t+1}\|^2 - \|w_t\|^2 }\\
= &\ \sqrt{\sum_{t \in I_1 \cup I_2} \|  w_{t} + y_t x_t\|^2 - \|w_t\|^2 }\\
= &\ \sqrt{\sum_{t \in I_1 \cup I_2} 2y_t w_t\cdot x_t - \|x_t\|^2} \\
\leq &\ \sqrt{\sum_{t \in I_1 \cup I_2} \|x_t\|^2}
\leq \sqrt{M'r^2}.
\end{align*}
This implies that $M' \leq \frac{r^2}{\rho^2}$. Since $M = M' + M_3$, a bound on $M_3$ is needed. We revert to a weaker bound and consider the expectation of
$M_3$. Since 
\[M_3 = \sum_{t=1}^T 1_{\left\{\widehat{y}_t = -1 \cap y_t = 1\right\}}1_{k_t = 0}, \]
we get that
\begin{align*}
\E[M_3] = & \sum_{t=1}^T \E[1_{\left\{\widehat{y_t} = -1 \cap y_t = 1\right\}}]\E[1_{k_t = 0}] \\
 = &\ (1 - p)\sum_{t=1}^T \E[1_{\left\{\widehat{y_t} = -1 \cap y_t = 1\right\}}],
\end{align*}
where the last equality uses the independence of $k_t$ from the data. To control the sum of expectations in this expression we consider the expectation of $M_2$. Since
\[
M_2 =  \sum_{t=1}^T 1_{\left\{\hat{y_t} = -1 \cap y_t = 1\right\}}1_{k_t = 1},
\]
then
\begin{align*}
\E[M_2] =&\ \sum_{t=1}^T \E[1_{\left\{\hat{y_t} = -1 \cap y_t = 1\right\}}]\E[1_{k_t = 1}] \\
=&\ p\sum_{t=1}^T \E[1_{\left\{\hat{y_t} = -1 \cap y_t = 1\right\}}].
\end{align*}
Thus
\[
E[M_3] = \ \frac{1-p}{p}\ \E[M_2].
\]
We can apply the crude bound $\E[M_3] \leq \frac{1-p}{p}\ \E[M']$, since $M_2 \leq M'$. Thus the expected number of mistakes can be bounded by
\begin{align*}
\E[M]  =  &\ \E[M'] + \E[M_3] \\
\leq & \ \frac{r^2}{\rho^2} + \frac{1-p}{p}\frac{r^2}{\rho^2} \\
\leq  & \ \frac{1}{p}\frac{r^2}{\rho^2}
\end{align*}
Note that this is a weak bound.  Not only is it greater than the
original perceptron bound by a factor of $\frac{1}{p}$, it is also only a
bound in expectation.

@@

\noindent Adaptive $p$ Perceptron algorithm:
\begin{quotation}
%\noindent Modified Perceptron:\\
\noindent $w_1=0$\\
\noindent $p=1$\\
{\tt for $t=1$ to $T$}:\\
\indent receive($x_t$)\\
\indent $\widehat{y}_t=\mbox{sign}(w_t\cdot x_t)$\\
\indent {\tt if} $\widehat{y}_t =1$ : \\
\indent\indent receive($y_t$)\\
\indent\indent {\tt if $y_t\neq\widehat{y}_t$}:\\
\indent\indent\indent $w_{t+1} = w_t + y_t x_t$\\
\indent\indent {\tt else $w_{t+1} = w_t$}\\
\indent {\tt else if ($\widehat{y}_t =-1$ {\tt and} $k_t = 1$)}:\\
\indent\indent {\tt if $y_t\neq\widehat{y}_t$}:\\
\indent\indent\indent $w_{t+1} = w_t + y_t x_t$\\
\indent\indent\indent $p=p + \gamma_1(1 - p)$\\
\indent\indent {\tt else }\\
\indent\indent\indent $w_{t+1} = w_t$\\
\indent\indent\indent $p = p - \gamma_2(p - p_{\min})$\\
\indent {\tt else if $\widehat{y}_t = -1$ and $k_t=0$}:\\
\indent\indent $w_{t+1} = w_t$\\
{\tt return $w_{T+1}$}
\end{quotation}
Note that $p_{\min}$, $\gamma_1$ and $\gamma_2$ are parameters that can be tuned. It is easy to extend the bound above from the modified Perceptron to the adaptive $p$ case.  Let $I_i$, $M_i$, and $M'$ be defined as above. The bound $M'\leq r^2/\rho^2$ follows in exactly the same way.  We then bound $M_3$ in expectation, 
\begin{align*}
\E[M_3] = & \sum_{t=1}^T \E[1_{\left\{\widehat{y_t} = -1 \cap y_t = 1\right\}}]\E[1_{k_t = 0}] \\
 \leq &\ (1 - p_{\min})\sum_{t=1}^T \E[1_{\left\{\widehat{y_t} = -1 \cap y_t = 1\right\}}],
\end{align*}
and bound $\E[M_2]$ from below
\begin{align*}
\E[M_2] =&\ \sum_{t=1}^T \E[1_{\left\{\hat{y_t} = -1 \cap y_t = 1\right\}}]\E[1_{k_t = 1}] \\
\geq &\ p_{\min}\sum_{t=1}^T \E[1_{\left\{\hat{y_t} = -1 \cap y_t = 1\right\}}].
\end{align*}
Continuing as before, this gives the following bound on $\E[M]$,
\begin{align*}
\E[M] = &\ \E[M'] + \E[M_3] \leq \frac{r^2}{\rho^2} + \frac{1-p_{\min}}{p_{\min}} \frac{r^2}{\rho^2}\\
= &\ \frac{1}{p_{\min}}\frac{r^2}{\rho^2}
\end{align*}
Unfortunately, we do not have a stonger theoretical bound for this adaptive $p$ approach, but we test this scheme in section \ref{sec:results} and see that it performs fairly well while requiring less input from the user.

@@

The files {\tt Perceptron.py} and {\tt OnlineSVM.py} house the classes that compose the different classifying algorithms we use.
The classes contain the methods that run the algorithm and the attributes such as weights and parameters that the algorithm uses and updates.
Within {\tt Perceptron.py} is the original Perceptron algorithm, a Margin Perceptron classifier, and a Kernel Perceptron classifier supplied with both a polynomial and Gaussian kernel.
Similarly, the Online SVM algorithm is implemented in the file {\tt OnlineSVM.py}.
Each algorithm takes a learning rate $\eta$ as a parameter. The modified Perceptron requires no extra parameters, while the Gaussian Kernel Perceptron classifier requires a $\sigma$ specifying the variance in the kernel.
The Polynomial Kernel Perceptron takes two parameters $c$ and $d$ which are the constants used in constructing the polynomial kernel.
The Margin Perceptron requires a parameter $\rho$ designating the separating margin of the feature vectors, and the Online SVM algorithm takes a parameter $C$ for the penalty applied to the update (i.e. $w_{t+1} = w_t -\eta(w_t- Cy_tx_t)$).
Each algorithm class has an associated wrapper that instantiates a class, and we use each of these algorithms when testing the tool.

@@

We tested the different online learning algorithms with sample data generated by graduate students at NYU.  Several students were asked to use {\tt getTestData.py} and {\tt LabelData.py} to gather the abstracts from Arxiv in categories that they specified and generate labels for each entry indicating whether they would be interested in reading a paper based on the abstract.\\

@@

It was interesting to see how varying $p$ the probability of presenting the paper had an impact on the results, some which might not have been entirely expected. For instance I would expect that asking for the most labels from the user (i.e. $p = 1$) would translate to considerably higher gains in both precision and recall all around. Yet for the best performing algorithms there wasn't a huge difference in recall between the adaptive version and $p = 1$.  For precision, the lower $p$ algorithms actually performed comparably to the original schemes.

\begin{figure}\centering
\includegraphics[width=0.7\textwidth]{EveryoneRecallPaper.pdf}
\includegraphics[width=0.7\textwidth]{EveryonePrecisionPaper.pdf}
\label{fig:PrecisionandRecall}
\caption{Performance of tested online classification algorithms using different values of $p$.  The adaptive $p$ algorithm performs relatively well compared to the original algorithm ($p = 1$), and requires less than half as many negative updates. The $p = 0.1$ case tends to perform poorly, but requires far fewer updates.  The majority vote classifier combines all others in this figure.}
\end{figure}

\section{Conclusions}
To summarize our work, we're pleased to have a developed a tool that we believe can be of use for many people across several fields. We use online learning algorithms to incrementally learn the user's preferences and suggest interesting articles daily.\\

In order for this approach to be useful, it cannot present the user with every single prediction.  Instead, we weaken the algorithms to present the user with all articles it predicts will be interesting, and only some of those that it predicts will not be interesting.  We show that in this context, a bound on the expected number of mistakes can still be achieved under the same assumptions as Theorem 7.8 from \cite{FML}.  Finally, we test several algorithms on labeled data, and show that they perform comparably to the original learning algorithms, but require less input from the user.  \\
In the future, it would be nice to develop tighter theoretical bounds on our modified learning algorithm, and particularly to develop an improved bound on the adaptive $p$ algorithm.  It is possible that such a bound would lead to a more informed decision on how to change $p$ in the algorithm to minimize mistakes. Perhaps giving a different functional form to the way $p$ is changed can produce an algorithm that minimizes the number of labels required from the user.\\
Finally, it would be worthwhile to explore other online learning algorithms and possibly other features that might improve predictive performance. Considering a wider variety of text classification models would probably produce better results, and the limited time to develop this tool compromised our ability to develop and play around with some of these models.
\bibliography{paper}
\end{document}

@@

%To summarize our work, we're pleased to have a developed a tool that we believe can be of use for many people across several fields and to have employed a number of machine learning features in its construction, from the text categorization model to the classification algorithms. Beyond constructing these features we created some of our own features. For instance the adaptive probability $p$ turned out to perform fairly well in comparison to the tool that didn't filter results based on preferences. It was also interesting to see the bound that we got adapted from the original Perceptron model. One point to mention about future work in this area is that it seems likely that some of the other bounds from the Perceptron algorithm can probably be applied to get similar bounds for this problem as well.

%Online learning methods are used to learn the user's preferences over time, but in order for them to be useful, they cannot present the user with every single prediction.  Instead, we weaken the algorithms to present the user with all articles it predicts will be interesting, and only some of those it feels will not be interesting.  We show that in this context, a bound on the expected number of mistakes can still be achieved under the some assumptions.  Finally, we test several algorithms on labeled data, and show that they perform comparably to the original learning algorithms, but require less input from the user.

%In the future, it would be nice to develop tighter theoretical bounds on our modified learning algorithm, and particularly to develop an improved bound on the adaptive $p$ algorithm.  It is possible that such a bound would lead to a more informed decision on how to change $p$ in the algorithm to minimize mistakes. Perhaps giving a different functional form to the way $p$ is changed can produce an algorithm that minimizes updates, though this guess is comes from our experience with the AdaBoost algorithm and its modifications.

%Finally, it would be worthwhile to explore other online learning algorithms and possibly other features that might improve predictive performance. Considering a wider variety of text classification models would probably produce better results, and the limited time to develop this tool compromised our ability to develop and play around with some of these models. Nevertheless that is a huge body of work in and of itself.
%
%
%
%
%
%
%
%
%
